---
title: "Practical Questions"
author: "YULIA"
date: "3/3/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

***PRACTICAL QUESTIONS***
Gun Control. 

```{r}
#installation_needed  <- TRUE
#loading_needed <- TRUE
#package_list <- c('foreign', 'xtable', 'plm','gmm', 'AER','stargazer','readstata13', 'boot', 'arm', 'lmtest', 'sem', 'bdynsys', 'ivpack')
#if(installation_needed){install.packages(package_list, repos='http://cran.us.r-project.org')}
#if(loading_needed){lapply(package_list, require, character.only = TRUE)}

# clear the global workspace

rm(list=ls())
library(AER)
library(plm)
data("Guns")
```

**3.1 PART 1**  
**For this part we will focus on the coefficients and not the standard errors**    
(you may use the default OLS standard errors assuming homoskedasticity for this part only).  

**1. Make a table that summarizes the basic properties of all the variables in the data set.**  

```{r}
mydata <- pdata.frame(Guns,index = c("state","year"), drop.index = FALSE)
#table(index(mydata), useNA = "ifany")
summary(mydata)
```
**2. Perform a pooled OLS analysis, ignoring the panel properties of the data set and just treating all the data as iid. Report the results.**  



Our main explanatory variable of interest is whether a state has a shall carry law, that is, a law that allows basically anyone to get a license to carry a gun around with them. So these are laws that encourage more guns on the street, not fewer. For every regression we will include the shall-carry law dummy variable as well as controls for the gender and racial makeup of the state's population, the population density, the state average income, and the states's prison population.  
```{r}
linear_model <- violent ~ law  + prisoners + male + density + cauc + afam + population + income
summary(linear_model)
#linear_model <- lm(violent ~. - murder -robbery - state - year, data = mydata)
#summary(linear_model)

naive_ols <- plm(linear_model, mydata, model ="pooling", index = c("state", "year"))
summary(naive_ols)
```

**3. Now perform the Feasible Generalized Least Squares analysis of this panel with "classical" random effects at the state level.**  
```{r}
RE_fit <- plm(linear_model, data = mydata, model ="random", index = c("state", "year"))
summary(RE_fit)
```

**4. Now perform a fixed effects analysis of the panel with fixed effects for the  state, estimated by both the "within" estimator and the "first differences" estimator.**     
Do you see any major difference in the two estimators here?  
Give some intuition for what you see.    


*Fixed Effects*
```{r}
fixed_effects_fit <- plm(linear_model, mydata, model="within", effect = "individual", index = c("state", "year"))
summary(fixed_effects_fit)
```


*First Differences*
```{r}
FD_fit <- plm(linear_model, mydata, model="fd", index = c("state", "year"))
summary(FD_fit)
```

We can observe that R-Squared of the FE regression is larger since it preserves more information. The coefficients on male and population differ a lot in FE and FD.  


**5. Now perform fixed effects analysis of the panel using the "within" estimator with state and year effects.**    
Perform first differences analysis adding in dummies for the year. Do you see any major difference in the two procedures? Explain. 


*Fixed Effects*
```{r}
twoways_fixed_effects_fit <- plm(linear_model, mydata , model="within", effect = "twoways", index = c("state", "year"))
summary(twoways_fixed_effects_fit)
```
*First Differences*
```{r}
FD_fit_year <- plm(violent ~ law  + prisoners + male + density + cauc + afam + population + income + factor(year), mydata, model="fd" , index = c("state", "year"))
summary(FD_fit_year)
```

Still large difference in coefficients is observed.  


**3.2 PART 2**  
**1. Compute the canonical White standard errors for this model and report the results.**  

```{r}
HC_coefs <- vcovHC(fixed_effects_fit, method = "white1")
white_ses_state <- sqrt(diag(HC_coefs))
print(white_ses_state)
```
**2. Now compute the canonical "clustered" standard errors with clustering at the state level (as in Arrelano 1987).**    

```{r}
HCV_coefs_state <- vcovHC(fixed_effects_fit, method = "arellano", cluster = "group")
clustered_ses_state <- sqrt(diag(HCV_coefs_state))
print(clustered_ses_state)
```

**3. Now compute the canonical "clustered" standard errors with clustering at the time unit (as in Arrelano 1987 but for T not N).**

```{r}
HCV_coefs_time <- vcovHC(fixed_effects_fit, method = "arellano", cluster = "time")
clustered_ses_time <- sqrt(diag(HCV_coefs_time))
print(clustered_ses_time)
```
**4. Now compute the Newey-West standard errors for panel data.**  

```{r}
NW_coefs <- vcovNW(fixed_effects_fit)
NW_ses_state <- sqrt(diag(NW_coefs))
print(NW_ses_state)
```
**5. Comment on the relative magnitude of these errors for the ?? on our key variable, the shall-carry law. Give some intuition for what you see.**  
Arellano approach gives highest standard errors and NW lowest. 


